---
title: "Binomial Functions"
author: "Beverly Leslie Tanada"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(binomial)
```

# Binomial Distribution

The Binomial distribution is perhaps the most famous probability distribution among discrete
random variables. This is the theoretical probability model that we use when calculating
probabilities about the number of successes in a fixed number of random trials performed
under identical conditions (assuming a constant probability of success on each trial). Below is the function:

$$Pr(X = k) = {n \choose k}p^k(1-p)^{n-k}$$
where:
• n is the number of (fixed) trials (n ≥ 0)
• p is the probability of success on each trial (0 ≤ p ≤ 1)
• 1 − p is the probability of failure on each trial
• k is a variable that represents the number of successes out of n trials (0 ≤ k ≤ n)

The package `"binomial"` will help to find various functions regarding the binomial distribution. The function `bin_variable()` will print the parameters: the number of `trials` and the `prob` of success.

```{r}
bin1 <- bin_variable(trials = 10, prob = 0.2)
bin1
```


The `summary` function gives you the list of parameters and a summary of the measurements.

```{r}
summary(bin1)
```

In alternative, we can also invoke individual elements of the summary through its own function.

### Mean and Variance

When X is a random variable that has a binomial distribution with n trials and probability of success p, we write X ∼ Bin(n, p). As with most random variables, it is useful to know about the various summary measures of X. The expected value or mean of a binomial distribution is: $$np$$. This is the expected number of successes in n trials. The variance is given by: $$np(1 − p)$$. Consequently, the standard deviation is simply the square root of the variance, that is: $$\sqrt{np(1 − p)}$$

The function `bin_mean()` will calculate the binomial distribution's mean.

```{r}
bin_mean(trials = 3, prob = 0.3)
```

While the function `bin_variance()` will calculate the binomial distribution's variance.

```{r}
bin_variance(trials = 3, prob = 3)
```

### Mode

For $$0 < p < 1$$, the most likely number of success in n independent trials with probability p of success on each trial is m, the greater integer less than or equal to $$np + p$$: 

$$m = int(np + p)$$

where int denotes the integer part function. If $$np + p$$ is an integer, as in the case p = 0.5 and
n odd, then there are two most likely numbers, m and m − 1. Otherwise, there is a unique
mode.

The function `bin_mode()` will calculate the binomial distribution's mode.

```{r}
bin_mode(trials = 3, prob = 0.3)
```

### Skewness and Kurtosis

Skewness is a measure of the asymmetry of the probability distribution of a random variable about its mean. The skewness value can be positive or negative, or undefined. The skewness of a binomial random variable can be calculated as:

$$skewness = \dfrac{1 − 2p}{\sqrt{np(1 − p)}}$$

The Kurtosis (from greek kurtos, meaning “curved, arching”) is a measure of the “tailedness”
of the probability distribution of a random variable. In a similar way to the concept of
skewness, kurtosis is a descriptor of the shape of a probability distribution. For a binomial
random variable, its kurtosis can be obtained as:

$$kurtosis = \dfrac{1 − 6p(1-p)}{np(1 − p)}$$

The function `bin_skewness()` will calculate the skewness of the binomial distribution.

```{r}
bin_skewness(trials = 3, prob = 0.3)
```

While the function `bin_kurtosis()` will calculate the kurtosis of the binomial distribution.

```{r}
bin_kurtosis(trials = 3, prob = 0.3)
```

## Finding number of combinations

Using the function `bin_choose()` will give the number of combinations in which `k` successes occur in `n` trials.

```{r}
bin_choose(n = 5, k = 2)
```

## The binomial probability

We can also calculate the probability of our binomial distribution - where we calculate the probability of getting a certain number successes in the number of desired trials. We use the `bin_probability()` function in this case.

```{r}
bin_probability(success = 0, trials = 5, prob = 0.5)
```

## The binomial distribution

Now, we can take it a step further and calculate the probability of success for each number of successes possible given the number of trials. We will use the `bin_distribution()` function.

```{r}
bin1 <- bin_distribution(trials = 5, prob = 0.5)
bin1
```

## Plot of binomial distribution

We can use the function `plot()` to see the histogram for the probability for each number of successes.

```{r}
plot(bin1)
```

## The binomial cumulative distribution

Lastly, we can find the binomial cumulative distribution where it calculates the cummulative probabilities from the previous number of successes. The function `bin_cumulative()` gives this result.

```{r}
bin2 <- bin_cumulative(trials = 5, prob = 0.5)
bin2
```

## Plot of binomial cumulative distribution

We can also plot the binomial cumulative distribution graph through invoking the `plot()` function.

```{r}
plot(bin2)
```

